---
title: "BIOST HW1"
author: "Shuyang Wu"
date: "1/29/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(boot)
library(ATE)
```

## Question 1

(a) The marginal distribution of quality of life: 

good: 0.44    moderate: 0.3   poor: 0.26

Conditional probability of qualiy of life given "high" grade histology: 

good: 0.118    moderate: 0.294    poor: 0.588

Conditional probability of qualiy of life given "low" grade histology: 

good: 0.606    moderate: 0.303    poor: 0.091

The conditional distributions of quality of life are different. This makes sense, because intuitively patients with more severe cancer prognosis are more likely to have a poor quality of life (0.588) comparing to patients with low grade histology (0.091). 

(b)
Missing completely at random means that whether or not participants answered the questions regarding their quality of life is completely random and independent of other events (for example histology). 

Missing at random means that the propensity for a data point to be missing is related to some of the observed data such as histology and not related to the quality of life answer itself.

Missing not at random means that the missing of entries for quality of life is dependent on other un-observed/un-measured variables such as patients' education level etc.

(c)
Table 2: Missing not at random -- all estimates are Biased
The marginal Distribution of quality of life: 

good: 0.605    moderate: 0.275   poor: 0.119

Conditional probability of qualiy of life given "high" grade histology: 

good: 0.231    moderate: 0.385    poor: 0.385

Conditional probability of qualiy of life given "low" grade histology: 

good: 0.723    moderate: 0.241    poor: 0.036

Table 3: Missing Completely at random -- all estimates are Unbiased
The marginal Distribution of quality of life: 

good: 0.44    moderate: 0.3   poor: 0.26

Conditional probability of qualiy of life given "high" grade histology: 

good: 0.118    moderate: 0.294    poor: 0.588

Conditional probability of qualiy of life given "low" grade histology: 

good: 0.606    moderate: 0.303    poor: 0.091

Table 4: Missing at random 
The marginal Distribution of quality of life: -- (biased) 

good: 0.481   moderate: 0.301   poor: 0.218

Conditional probability of qualiy of life given "high" grade histology: -- (unbiased)

good: 0.118    moderate: 0.294    poor: 0.588

Conditional probability of qualiy of life given "low" grade histology: --(unbiased)

good: 0.606    moderate: 0.303    poor: 0.091

(d)
```{r Table2}
#good
(60*132/(132-49) + 6*68/(68-42))/200
#moderate
(20*132/(132-49) + 10*68/(68-42))/200
#poor
(3*132/(132-49) + 10*68/(68-42))/200
```
Table 2
The marginal distribution of quality of life: 

good: 0.555    moderate: 0.290   poor: 0.155

```{r Table3}
#good
(60*132/(132-33) + 6*68/(68-17))/200
#moderate
(30*132/(132-33) + 15*68/(68-17))/200
#poor
(9*132/(132-33) + 30*68/(68-17))/200
```
Table 3: 
The marginal Distribution of quality of life: 

good: 0.44    moderate: 0.3   poor: 0.26

```{r Table4}
#good
(60*132/(132-33) + 4*68/(68-34))/200
#moderate
(30*132/(132-33) + 10*68/(68-34))/200
#poor
(9*132/(132-33) + 20*68/(68-34))/200
```
Table 3: 
The marginal Distribution of quality of life: 

good: 0.44    moderate: 0.3   poor: 0.26

## Question 2
```{r mental_baseline}
mental_baseline <- read.csv("~/Downloads/mental_baseline.csv")
mental_baseline$M <- rep(0, 500)
mental_baseline$M[is.na(mental_baseline$sat)] <- 1

#look at missing pattern
plot(mental_baseline$hospday,mental_baseline$M)
plot(mental_baseline$cprs,mental_baseline$M)

#check for dependencies and calculate propensity scores
propensity <- glm(M ~ cprs + hospday, family=binomial, data=mental_baseline)
summary(propensity)
ps <- propensity$fitted.values

#complete case analysis
cc <- na.omit(mental_baseline$sat)
mean(cc)

#IPW
weighted.mean(mental_baseline$sat, 1/ps, na.rm = T)

ipw.mean=function(data1,i){
  fit.ps=glm(M~ hospday + cprs,data=data1[i,],family=binomial)
  tmp.ps=fit.ps$fitted.value
  weighted.mean(data1$sat[i],1/tmp.ps,na.rm=T) 
  
}

ipw.mean(mental_baseline, 1:nrow(mental_baseline))

boot.out = boot(mental_baseline, ipw.mean, R = 100)

boot.ci(boot.out, index = 1, type = c("norm", "perc"))
```
(a) Missing data pattern: Missing Completely at Random
(b) Based on the logistic regression model output, hospday has a little significant effect with on the missingness of sat, cprs has none. I think the missingness mechanism is missing completely at random.
(c) Complete case analysis result: 19.1037
    IPW: 18.92381
    95% confidence interval: (18.41, 19.37)
(d) The two estimate are similar because the missing pattern is Missing Completely at Random, meaning that the missingness is not dependent on any observed variables but itself. Therefore, using both inverse probability weighting and complete case analysis shouldn't make a difference on the results obtained.

## Question 3

```{r bc}
#(a)
bc <- read.csv("~/Downloads/bc.csv")
propensity <- glm(treatment ~ tgr + age, family=binomial, data=bc)
summary(propensity)

#(b) treatment effect
#empirical mean difference
mean(bc$tgr) - mean(bc$tmass)

#IPW
ipw.ATE=function(data1,i){
  fit.ps=glm(treatment ~ tgr + age,data=data1[i,],family=binomial)
  tmp.ps=fit.ps$fitted.value
  mean(data1[i,2]*data1[i,1]/tmp.ps)-mean((1-data1[i,2])*data1[i,1]/(1-tmp.ps)) #mean difference (E(Y1)-E(Y2))
}

ipw.ATE(bc,1:nrow(bc))

#non-parametric calibration
Y<-bc[,1]    
treat<-bc[,2]  
X<-as.matrix(bc[,-c(1,2)]) 
fit1<- ATE(Y,treat,X)
fit1
summary(fit1)
```

(c) empirical mean difference: 13.0031
    IPW: 0.04724655
    non-parametric calibration: 0.09406
    Because the treatment assignment is largely dependent on age, empirical mean difference does not take into account the dependency, its value is relatively larger than those from IPW and non-parametric calibration. 
    
##Question 4
(a) I have a dataset at hand, it includes head and neck cancer patients' clinical information, but some of the data were undetermined, for example, tumor stage, HPV status, and metastastic levels. The missingness for tumor stage is missing completely at random; Missing at random or possibily not at random for metastastic levels and HPV status.
